Titeln på avhandlingens språk

### 2023-05-10 body/acks.tex
acks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ACKNOWLEDGMENT %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Acknowledgements}
\glsresetall%

I would like to begin by thanking my supervisor, Prof. James Gross, for his guidance and patience during this long process.\todo{todo}

I would like to express my deepest gratitude to Prof. Mahadev ``Satya'' Satyanarayanan, for hosting me repeatedly, over long periods of time, at \gls{CMU} as a visiting researcher.
His tremendous experience and generosity have been invaluable, and his guidance and advise have certainly shaped my work.
Likewise, I'm deeply indebted to Prof. Roberta ``Bobby'' Klatzky of \gls{CMU} for the many helpful and insightful discussions, for introducing me to the field of Human-Computer Interaction, and for always greeting me with a warm smile.

I would also like to thank everybody involved in the defense of this thesis.
I am very grateful to Prof. Yu Xiao of Aalto University, for agreeing to act as my dissertation opponent.
My dissertation defence will certainly benefit greatly from her expertise and insights.
I would like to extend my sincere thanks to my grading committe, consisting of Prof. Ana Aguiar (Universidade do Porto), Prof. Per Gunningberg (Uppsala Universitet), and Prof. Klaus Wehrle (\gls{RWTH}), for investing time and effort into reading and grading this dissertation.
Special thanks also go to Prof. Markus Flierl for advance-reviewing this dissertation, and to Prof. Mats Bengtsson for agreeing to act as chairperson during the defense.

I also extend my thanks to my colleagues and co-authors.
To Samie, Vishnu, and Neel at \pgls{KTH}, for always being willing to engage in interesting discussions, but also for generally being fun to be around.
And to Junjue and Babu Pillai from \gls{CMU} for enriching my research with their interesting insights.

I am very lucky and grateful that I got to stay at both \pgls{KTH} and \gls{CMU} over the years, and I would like to thank everyone who made these spaces welcoming and engaging.
A big thank you goes out to my colleagues (both current and former) at the division of \gls{ISE} for all those fun lunches at Restaurang Q, Cypern, or SysterOBror over the years, which always helped to break the monotony of the day.
This includes but is not limited to Samie, Vishnu, Neel, Sahar, Antonios, Germán, Boules, Baptiste, Håkan, Pol, Marie, and Sebastian.
I would additionally like to extend a special thank you to Sebastian Schiessl, for ``taking me under his wing'' when I first arrived at \gls{ISE} and acting as a sort of mentor during my first moths.
Another big thank you goes out to everyone (current and former) at Satya's group at \gls{CMU}, for always welcoming me with open arms and treating me like one of them whenever I was visiting.
This includes but is not limited to Junjue, Roger, Shilpa, Tom, Jan, Edmond, Jim, Babu, and Zhuo.

I would like to extend my sincere thanks to Prof. Sandra Céspedes, of Universidad de Chile and Concordia University, for her guidance and urging to embark on this journey towards a Ph.D., and for always being willing to offer help and advice.
Likewise, I'd like to thank Prof. Javier Bustos, of Universidad de Chile and NIC Chile Research Labs, for giving me an opportunity many years ago to get into distributed systems research at NIC Labs.
Also, I'd like to extend my gratitude to Prof. José Miguel Piquer, of Universidad de Chile, for taking me as his ``long-distance'' teaching assistant and allowing me to give back to my \emph{alma mater} from across the ocean.

@@

This dissertation would not have been possible without the help and support of my family.
First and foremost, I would like to thank the love of my life and wife~\footnote{\emph{future-wife} at the time of writing, but \emph{wife} at the time of the defense of this thesis}, Andresa.
I could not have done this without her companionship, love, and unconditional support.\todo{TODO}

I am also deeply grateful to my parents, Gabriel and Valeria, for always encouraging my curiosity and my drive to learn and discover.
They have been my rocks, my pillars, since the day of my birth, and I am proud to be their son.
A big thank you goes out to my sister, Paola, as well.
Although we sometimes get on each other's nerves, she has always been there for me and supported me, and I could not have asked for a better sister.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ACKNOWLEDGMENT %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Acknowledgements}
\epigraph{La educación es, tal vez, la forma más alta de buscar a Dios.}{\emph{Gabriela Mistral}}
\glsresetall%

There are many, many people I owe gratitude to for making this dissertation possible.
I would like to begin by thanking my supervisor, Prof. James Gross.
This dissertation would not have been possible without his guidance and motivation during this long process.
His supervision opened doors that I had never imagined would open for me, and his dry German humor has certainly helped cheer me up whenever I was overwhelmed by some deadline.
Above all, I would like to thank James for his patience and understanding whenever I encountered difficulties, and for the flexibility he has afforded me in completing this Ph.D.

I would like to express my deepest gratitude to Prof. Mahadev ``Satya'' Satyanarayanan, for welcoming me repeatedly, over long periods of time, as a visiting researcher at \gls{CMU}.
Likewise, I'm deeply indebted to Prof. Roberta ``Bobby'' Klatzky of \gls{CMU} for the many helpful and insightful discussions, for introducing me to the field of Human-Computer Interaction, and for always greeting me with a warm smile.
Both Satya's and Bobby's tremendous experience and generosity have been invaluable, and their guidance and advice have certainly shaped my work for the better.

I would also like to thank everybody involved with the defense of this thesis.
I am very grateful to Prof. Yu Xiao of Aalto University, for agreeing to act as my dissertation opponent.
My dissertation defense will certainly benefit greatly from her expertise and insights.
I would like to extend my sincere thanks to my grading committe, consisting of Prof. Ana Aguiar (Universidade do Porto), Prof. Per Gunningberg (Uppsala Universitet), and Prof. Klaus Wehrle (\gls{RWTH}), for investing time and effort into reading and grading this dissertation.
Special thanks also go to Prof. Markus Flierl of \pgls{KTH} for advance-reviewing this dissertation, and to Prof. Mats Bengtsson, also of \pgls{KTH}, for agreeing to act as chairperson for the defense.

Moreover, I extend my thanks to my co-authors and colleagues.
To Samie Mostafavi, Vishnu N. Moothedath, and Neelabhro Roy at \pgls{KTH}, for always being willing to engage in interesting discussions, but also for generally being fun to be around.
And to Dr. Jaya Prakash (IMDEA Networks Institute), Dr. Junjue Wang (\gls{CMU}), and Dr. Padmanabhan ``Babu'' Pillai (\gls{CMU}) for enriching my research with their interesting insights.

I am very lucky and grateful that I got to stay at both \pgls{KTH} and \gls{CMU}, and I would like to thank everyone who made these spaces welcoming and engaging.
A big thank you goes out to my colleagues (both current and former) at the division of \gls{ISE} for all those fun lunches at Restaurang Q, Cypern, and SysterOBror over the years, which always helped to break the monotony of the days.
This includes but is not limited to Samie, Vishnu, Neel, Sahar, Antonios, Germán, Boules, Baptiste, Håkan, Pol, Marie, Wendi, and Sebastian.
I would additionally like to extend a special thank you to Dr. Sebastian Schiessl, for ``taking me under his wing'' when I first arrived at \gls{ISE} and acting as a sort of mentor during my first moths.
Likewise, I extend my gratitute to Gerd Fransson, for greeting me with a huge smile every morning the years our offices faced each other and cheering up my days with her wholesome energy.
I also thank Prof. Joakim Jaldén for giving me the opportunity to act as his teaching assistant.
Another big thank you goes out to everyone (current and former) at Satya's group at \gls{CMU}, for always welcoming me with open arms and treating me like one of them whenever I was visiting.
This includes but is not limited to Junjue, Tom, Roger, Shilpa, Jan, Edmond, Jim, Babu, and Zhuo.
I will always remember our Monday morning talks and Friday lunch outings dearly.

Furthermore, I would like to extend my sincere thanks to Prof. Sandra Céspedes, of Universidad de Chile and Concordia University, for her guidance and urging to embark on this journey towards a Ph.D., and for always being willing to offer help and advice.
Likewise, I'd like to thank Prof. Javier Bustos, of Universidad de Chile and NIC Chile Research Labs, for giving me the opportunity many years ago to first get into distributed systems research.
Also, I'd like to extend my gratitude to Prof. José Miguel Piquer, of Universidad de Chile, for taking me as his ``long-distance'' teaching assistant and allowing me to give back to my \emph{alma mater} from across the ocean.

@@

This dissertation would not have been possible without the help and support of my family.
First and foremost, I would like to thank the love of my life and wife~\footnote{\emph{Future-wife} at the time of writing, but \emph{wife} at the time of the defense of this thesis!}, Andresa.
I could not have done this without her companionship and love.
I am eternally grateful for her unconditional support over difficult periods of time and over tremendous distances while we were living apart, and for making every day we spend together amazing.

I am also deeply grateful to my parents, Gabriel and Valeria, for always encouraging my curiosity and my drive to learn and discover.
They have been my rocks, my pillars, since the day I was born, and I am proud to be their son.
A big thank you goes out to my sister, Paola, as well.
Although we sometimes get on each other's nerves, she has always been there for me and supported me, and I could not have asked for a better sister.
Likewise, I extend my deepest gratitude to everyone in my immediate and extended family who have given me support and love over the years, including but not limited to my paternal grandparents, Manuel and María Yolanda, my maternal grandparents, Caupolicán and María Aura, my uncle Mauricio, his wife Marylen, their daughter Amanda, and everybody else.
I would like to make a special mention to my late grandfather, Caupolicán ``Poli'' Muñoz, who always inspired and fanned the flames of academic curiosity in me.
I hope he would have been proud seeing how far I have come.
Another special mention goes to my uncle Manuel, who graciously took me in and gave me a place to stay at the first few months of my Ph.D. studies.

I would also like to thank all of my friends who have made this long, ardous journey more amenable.
I would particularly like to thank everyone I met through Tartan Salsa, including, but not limited to, Andresa, Jennifer, Alyson, Kuai-Kuai, Carmen, Moataz and Nora, Eric, and many others.
They made Pittsburgh feel like home, even before I finally moved there permanently in 2022.
I also want to thank Ann-Charlotte in Stockholm, for always being so kind, so considerate, and so much fun to hang with.
Further, I am deeply grateful for my friends in Chile who have remained a source of support, even though we only get to see each other once a year when I visit.
This includes, but is not limited to, Juan Pablo, Fernanda, Diego, Camila, George, Patricio, Kevin, Pablo, and many others.
Yet another special mention goes to Diego's family, who have always treated me as one of their own.

Finally, I would be remiss in not mentioning my pets and animal friends who have provided me with love and warmth during these years.
I thank my dogs, Aros and Skatt, who I sadly had to leave behind in Chile with my sister, but who still greet me with enthusiastically wagging tails whenever I visit.
And I thank my cats, Mochi and Tiramisú, who, although have only been in my life for about a year, have sweetened and made much more bearable the long days and nights I have spent writing this dissertation.

\bigskip

\noindent{}Manuel Olguín Muñoz\\
Pittsburgh, May 2023
